---
title: "RWorksheet_Celestra#4b"
author: "Kenneth Celestra"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.
```{r}
vectorA <- c(1, 2, 3, 4, 5)

matrixA <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrixA[i, j] <- abs(i-j)
  }
}

print(matrixA)

```
2.
```{r}
# Number of rows
num_rows <- 5


for (i in 1:num_rows) 
  cat(strrep(' "*"', i), "\n")


```
3.
```{r}
# start <- as.integer(readline(prompt = "Enter the starting number for Fibonacci sequence: "))

start <- 1
a <- start
b <- 1

repeat {
  if (a >= start) {
    cat(a, " ")
  }
  
  next_number <- a + b
  a <- b
  b <- next_number
  
  if (a > 500) {
    break
  }
}
```
4.
a
```{r}
data <- read.csv("Shoe_sizes.csv")
head(data)
```
b
```{r}
maledata <- subset(data, Gender == "M")
femaledata <- subset(data, Gender == "F")

num_male <- nrow(maledata)
num_female <- nrow(femaledata)

cat("Number of Male observations:", num_male, "\n")
cat("Number of Female observations:", num_female, "\n")

```
c
```{r}
# Sample data for Household Data with counts of males and females
gender_counts <- table(data$Gender)

barplot(
  gender_counts,
  main = "Number of Males and Females", 
  col = c("blue", "pink"), # Colors for bars
  names.arg = c("Male", "Female"),
  xlab = "Gender",
  ylab = "Count",
)
legend(
  "topright",
  legend = c("Male", "Female"),
  fill = c("blue", "pink")
)

```
5.
a
```{r}
expenses <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)

percent_labels <- paste(names(expenses), round(expenses / sum(expenses) * 100, 1), "%")

pie(
  expenses,
  labels = percent_labels,
  col = c("blue", "green", "orange", "pink"),
  main = "Dela Cruz Family Monthly Expenses"
)

```
6.
```{r}
data(iris)
```
a
```{r}
str(iris)
```
The Sepal.Length representing the length of the sepal (in centimeters).
Sepal.Width representing the width of the sepal.
Petal.Length representing the length of the petal.
Petal.Width representing the width of the petal.
Species A factor variable with 3 levels: "setosa", "versicolor", and "virginica", representing three species of iris flowers.

b
```{r}
mean_values <- colMeans(iris[, 1:4])  

mean_values

```
c
```{r}
species_counts <- table(iris$Species)

percent_labels <- paste(names(species_counts), round(species_counts / sum(species_counts) * 100, 1), "%")

pie(
  species_counts,
  labels = percent_labels,
  col = c("lightblue", "lightgreen", "lightcoral"), # Colors for each species
  main = "Species Distribution in the Iris Dataset" # Title
)

legend(
  "topright",
  legend = names(species_counts),
  fill = c("lightblue", "lightgreen", "lightcoral")
)

```
d
```{r}
setosa_subset <- subset(iris, Species == "setosa")
versicolor_subset <- subset(iris, Species == "versicolor")
virginica_subset <- subset(iris, Species == "virginica")

print(tail(setosa_subset, 6))

print(tail(versicolor_subset, 6))

print(tail(virginica_subset, 6))

```
e
```{r}
iris$Species <- as.factor(iris$Species)

plot(
  iris$Sepal.Length,
  iris$Sepal.Width,
  main = "Iris Dataset", 
  sub = "Sepal width and length", 
  xlab = "Sepal Length (cm)",
  ylab = "Sepal Width (cm)", 
  pch = 21,
  col = iris$Species,
  bg = iris$Species, 
  cex = 1.5 
)
```
f
The scatterplot depicts the relationship between Sepal.Length and Sepal.Width for three iris species: setosa, versicolor, and virginica. Setosa is clearly clustered in the lower left quadrant, with smaller dimensions, distinguishing it from the other species. Versicolor exhibits a broader range of sizes, overlapping with virginica, which generally has larger measurements.
