---
title: "RWorksheet_Celestra#2."
author: "Kenneth Celestra"
date: "2024-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using Vectors
1.
a. It outputs a set of numbers ranging from -5 to 5.
```{r}
x <- -5:5
x
```
b. The value of x are numbers from 1 to 7
```{r}
x <- 1:7
```
2. 
a. It outputs a set of numbers starting from 1 and it gets incremented by 0.2 until it reaches 3.
```{r}
y <- seq(1,3, by=0.2)
y
```
3. 
```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
```
a. 22
```{r}
ages[3]
```
b. 28, 36
```{r}
ages[c(2,4)]
```
c.
```{r}
ages[-c(4,12)]
```
4. 
a. It outputs the vectors with their name
```{r}
 x[c("first", "third")]
```

b.
```{r}
x <- c("first"=3, "second"=0, "third"=9)
names(x)
```

5.
a&b. This changes the 2nd element from -2 into 0
```{r}
x <- c(-3:2)
x
```
```{r}
x[2] <- 0
```
6.
a.
```{r}
mrCruzdata <- data.frame(
  Month = c("Jan", "Feb", "Mar", "Apr", "May", "Jun"),
  Price_Per_Liter = c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00),
  Purchase_Quantity = c(25, 30, 40, 50, 10, 45)
)
mrCruzdata
```
b. 
```{r}
avr_fuel <- weighted.mean(mrCruzdata$Price_Per_Liter, mrCruzdata$Purchase_Quantity)
print(paste("Mr. Cruz avarage fuel expenditure is", avr_fuel))
```
7.
a. This would be the output if "rivers" were typed in the R Console:
```{r}
rivers
```
b&c.
The result would be that the data vector will access the rivers data sets and get its length, sum, mean, median, variance, standard deviation, its minimun and maximum.
```{r}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers),
sd(rivers), min(rivers), max(rivers))
data
```
8.
a.
```{r}
power_ranking <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)

celebrity_name <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", "David Letterman", "Phil Mickelson", "J.K. Rowling", "Bradd Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Leno", "Celine Dion", "Kobe Bryant")

pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)

power_ranking
celebrity_name
pay
```
b.
```{r}
celebrity_name[19] <- "J.K. Rowling"
power_ranking[19] <- 15
pay[19] <- 90

power_ranking
celebrity_name
pay
```
c.
```{r}
forbes_data <- read.csv("/cloud/project/Worksheet2/PowerRanking.csv")
```
d.
```{r}
ranks <- forbes_data[10:20, ]
save(ranks, file = "/cloud/project/Worksheet2/Ranks.RData")
```
e.
The output table containing rows 10 to 20. This includes the celebrities ranked from 10th to 20th position, their names, and their corresponding pay, which is stored in the csv file.



9.
a.
```{r}
library("openxlsx")
library("readxl")
vienna <- read_xlsx("/cloud/project/Worksheet2/hotels-vienna.xlsx")

```
b.
```{r}
dim(vienna)
```
c.
```{r}
selcol <- vienna[, c("country", "neighbourhood", "price", "stars", "accommodation_type", "rating")]
```
d.
```{r}
save(selcol, file = "/cloud/project/Worksheet2/new.RData")
```
e.
```{r}
load("new.RData")
head(selcol)
tail(selcol)
```
10.
a.
```{r}
veg <- list("Kangkong", "Papaya", "Sayote", "Okra", "Kalabasa", "Pechay", "Talong", "Ampalaya", "Sitaw", "Malunggay")
veg
```
b.
```{r}
veg <- append(veg, list("Labanos", "Patatas"))
veg
```
c.
```{r}
veg <- append(veg, list("Tawgi", "Lettuce", "Repolyo", "Pipino"), after = 5)
veg
```
d.
```{r}
veg <- veg[-c(5, 10, 15)]
veg
```







